Automatic Hadith Graphs using Arabic Morphological Analysis

### 2011-08-01 tex/hadith.ijcnlp.tex
some changes in tex file
corrected bugs in satemachine of hadith

corrected bugs in mergeGraphs and buildGraph using hash

The morphological analysis of the word \RL{.hd_tnA} 
``narrated to us'' that starts the hadith in Figure~\ref{f:exhadith}
returns the stem \RL{.hd_t}.
The stem  \RL{.hd_t} is also the stem of other 
words such as \RL{.hd_tny} and \RL{.hd_thm}.
The stem also shares similar part of speech (POS)
and English meaning gloss annotations with words
such as \RL{qaal} ``said'' and \RL{'a_hbar} ``told''. 
The use of morphological analysis allows us to 
subsume listing these stems and all of their morphological
derivations with one rule that matches the gloss and 
POS tags. 
Similarly the words \RL{ibn} ``the son of'' and \RL{bn} 
return the same morphological analysis.

@@

{\bf Sarf.~~~}
In this paper, 
we use Sarf~\footnote{Sarf (name modified for blind review) }
an {\em application-specific efficient
morphological analyzer} that uses 
non-deterministic finite state Automata 
driven by an application-specific controller customized 
by the user. 
%Each machine in Sarf takes one letter at a time as input
%and represents a valid morphological analysis.
The application-specific controller prunes false positives
early in the run by making a decision at every input letter
to control the machines.
Sarf builds on and extends the lexicon of 
Buckwalter~\shortcite{Buckwalter:02} 
with proper and location names extracted from different online 
sources~\footnote{\href{http://alasmaa.net/}{http://alasmaa.net/ }, 
\href{http://ar.wikipedia.org/}{http://ar.wikipedia.org/}}
%as well as biblical sources~\footnote{Genesis 4:17-23; 5:1-32; 9:28-10:32; 11:10-32; 25:1-4, 12-18; 36:1-37:2; Exodus 6:14-25; Ruth 4:18-22; 1 Samuel 14:49-51; 1 Chronicles 1:1-9:44; 14:3-7; 24:1; 25:1-27:22; Nehemiah 12:8-26; Matthew 1:1-16; Luke 3:23-38}.
as well as biblical sources.
%~\footnote{Genesis 4:17-23; 5:1-32; 9:28-10:32; 11:10-32; 25:1-4, 12-18; 36:1-37:2; Exodus 6:14-25; Ruth 4:18-22; 1 Samuel 14:49-51; 1 Chronicles 1:1-9:44; 14:3-7; 24:1; 25:1-27:22; Nehemiah 12:8-26; Matthew 1:1-16; Luke 3:23-38}.

@@

@@

The process also benefits from name connector 
words and narrator connector words to enhance the
accuracy of the chain extraction and to segment the 
chain accurately into narrators.
Once we meet a name connector or a narrator 
connector word we reset the word counters that we
compare against the thresholds. 
that directly interacted with Sarf.

@@

\section{The narrator graph}

@@

The directed acyclic graph (DAG) 
in Figure~\ref{f:narrators} shows a partial order relation (POR) 
between narrators.
We automatically extracted the POR from the three books of 
hadith~\cite{IbnHanbal,AlKulayni,AlTousi}.
The nodes in boxes are the \RL{matn} of the hadith, 
and the other nodes are the narrators.
Given a set of chains of narrators similar to 
Figure~\ref{f:exhadith} we formed the DAG using graph algorithms 
that merged equal names in the chains. 
We compared the narrators together using a morphological
distance function. 
We considered each narrator to have several morphological 
components and we rewarded every morphological match with 
a small $\delta$. 
Then we applied a threshold $\theta$ to the sum of rewards
and decided the equality of the narrators. 
%We omitted the details of the distance function for brevity.

\begin{table}[tb]
%\centering
%\resizebox{.9\columnwidth}{!}{
\begin{tabular} {p{6cm}}
\begin{Verbatim}[fontsize=\relsize{-2},
frame=topline,framesep=4mm,label=\fbox{MergeChains algorithm},
commandchars=\\\{\}, codes={\catcode`$=3\catcode`_=8}]
Merge(Chain $C$[])
  Chain $C_c$[]=\{\}
  foreach Chain $c_1$ in $C$ 
    $C_c = C_c \cup c_1$
    foreach Narrator $n_1$ in $c_1$
      foreach Chain $c_2$ in $C-C_c$
        foreach narrator $n_2$ in $c_2$

    \}
  \}
\end{Verbatim}
\end{tabular}
%}
\label{a:merge}
\end{table}

@@

@@

\subsection{Use of the POR graph}

inconsistency check 
consistency checks
annotate with information extracted from biographies
what if analysis

@@

\begin{table}[bt]
\centering
\caption{Results of the hadith case study with Sarf.}
%\begin{tabular}{|p{1.5cm}||c|c||c|c||c|c|} \hline
\resizebox{1.1\columnwidth}{!}{
\begin{tabular}{lcp{.2cm}cp{.2cm}c} %\cline{2-10}
 &  AlKafi & & AlIstibsar & &IbnHanbal \\ \cline{1-6}
Word count  &98,943 & & 103,835 & & 20,354 \\ 
 Names       & 12,060  & & 14,613& & 3,013\\
%Names/Narrator & 1.97 & & 1.84& & 1.25 \\
Narrators  & 2,623 & & 5,767& & 1,755 \\ 
%Narrators/Chain  & 4.84 & & 4.76 & &4.05 \\
Chains  & 542 &  & 1,211& & 433 \\ 
Ignored names  & 6,400 &  & 3,348 & & 642 \\ \hline
Segmentation accuracy  & 96\%& & 96\%& & 92\%\\ 
Chain accuracy & 99\%&  & 99\%& & 97\% \\ 
%Narrator accuracy  & & & & &  & & & \\ 
Narrator accuracy  & 91\%& & 90\% & & 90\% \\ \hline
%Name false positives  & 7\%&  & 4\% & & 4\% \\ \hline
Running time (secs.)& 1.32 & & 1.31 & & .096\\ \hline 
\end{tabular}
}
\normalsize
\label{t:hadithresallresults}
\end{table}

We ran our experiments using a dual core 2.66 Ghz 64-bit processor 
with 4GB of memory using a Linux operating system. 
We report our results in Table~\ref{t:hadithresallresults} on the 
three hadith books.
NaGEMA detected narrator chains and segmented the books 
correctly with a chain accuracy above 97\% and a narrator name
accuracy above 90\%
in all cases and across all the 1986 traditions in the three 
different books with different formats.

The row ``word count'' is an indicator of the size of the data.
The %numbers in 
rows ``names'', ``narrators'', and ``chains'' denote the total
number of the corresponding entities per book.
The row ``ignored names'' reports the number of names detected 
outside the chains and thus are irrelevant to
the application. 
%In addition, the rows ``names/narrator'' and ``narrators/chain'' 
%report the average
%number of names per narrator and 
%the average number of narrators per chain respectively.

The ``segmentation accuracy'' reports on the accuracy of 
segmentation of the book into separate hadiths.
We computed the metric manually by marking the beginning 
and end of a representative sample of hadiths in each 
book and compared that to the NaGEMA results. 
We used the same approach to compute all other accuracy measures.
The ``chain accuracy'' reflects the percentage of correct narrators 
covered by extracted chains.
The ``narrator name accuracy'' row reflects the percentage of 
correct proper names detected within
narrator structures.

%The row ``name false positives'' reports the percentage of words 
%recognized as names while they should not.
%We notice that this metric did not affect the accuracy
%of narrator and chain segmentation since detecting
%a narrator or a chain with the refined analysis without substantially affecting the 
%high abstract measures such as chain and segmentation
%accuracy.

The closest to our work is e-Narrator~\cite{Azmi2010}, a 
tool that uses a grammar based approach to solve 
the narrator extraction problem. 
The e-Narrator tool preprocesses the text to remove 
punctuation and diacritics and to normalize white
spaces. 
Then e-Narrator uses a similarity-based approach
to memory based learning~\cite{Azmi2010} in order to perform 
shallow parsing of the preprocessed text. 
The e-Narrator tool uses a parser based on a context
free grammar that enumerates several ways to express
the name of the prophet, several possible words that 
might precede the name of the prophet, several possible
words that might separate two narrators, and several
words that might separate parts of the name of the narrators.
The e-Narrator tool fails when a variation of one of the 
listed words and not an exact form of it
happens in the text. 
In their grammar they consider a name to be any sequence of 
legal Arabic characters. 
The e-Narrator tool generates successfully 86.7 percent of
the narration chains of 90 selected traditions with 
34 simple cases and 56 hard cases.

NaGEMA does not need a preprocessing step since 
it uses morphological analysis to process the text. 
While e-Narrator bases its analysis on the patterns 
of words that surround the narrator names, 
NaGEMA bases its analysis on the fact that narrator
names are likely to happen in localities with a high
number of proper names. 
NaGEMA uses the narrator connectors as helpers and
uses morphological analysis to compare the stems of
those words and thus NaGEMA avoids the problem of 
``parser noise words'' that e-Narrator faces. 
Even though we do not have access to the selected
90 traditions e-Narrator reports on, we claim
that NaGEMA outperforms e-Narrator as it works on
the full text of the presented hadith books
and reports a higher success rate.

@@

We presented NaGEMA, a narrator chain extractor
using morphological analysis. 
NaGEMA extracted narrator chains with above 90\% accuracy
from three different hadith books and generated 
directed acyclic graphs relating the narrators 
in a partial order relation.

This partial order graph is instrumental to automate
localization of narrators in biography documents and
segmenting biography documents in the future.
Given that a biography discusses a narrator 
and mentions his professors and students,
we plan to segment the biographies with a graph 
coloring algorithm
that traverses the text and colors the POR node whenever
a name is found.

Once the biographies are analyzed, we can annotate
the POR with qualifiers of the narrators that reflect
their authenticity. 
%We can also annotate the POR with the locations and 
%the time the narrators lived in. 
%Then we can compute a time and location overlap
%check to discover inconsistent narrations.
%We can also perform several interesting checks using 
%this POR on its own such as checking for the effect of
%one narrator on the hadith literature.

We also plan to 
enhance the accuracy of hadith detection and the 
narrator merging algorithms.

>>>>

The morphological analysis of the word \RL{.hd_tnA} 
``narrated to us'' that starts the hadith in Figure~\ref{f:exhadith}
returns the stem \RL{.hd_t} ``narrated''.
The stem  \RL{.hd_t} is also the stem of other 
words such as \RL{.hd_tny} ``narrated to me'' and \RL{.hd_thm} ``narrated to them''.
The stem also shares similar part of speech (POS)
and English meaning gloss annotations with words
such as \RL{qaal} ``said'' and \RL{'a_hbar} ``told''. 
The use of morphological analysis allows us to 
subsume listing these stems and all of their morphological
derivations with one rule that matches the gloss and 
POS tags. 
Similarly the words \RL{ibn} ``the son of'' and \RL{bn} 
return the same morphological analysis.

@@

{\bf Sarf.~~~}
Sarf~\footnote{Sarf (name modified for blind review) }
is 
an {\em application-specific efficient
morphological analyzer} based on
non-deterministic finite state Automata 
driven by an application-specific controller customized 
by the user. 
%Each machine in Sarf takes one letter at a time as input
%and represents a valid morphological analysis.
The application-specific controller prunes false positives
early in the run by making a decision at every input letter
to control the machines.
Sarf builds on and extends the lexicon of 
Buckwalter~\shortcite{Buckwalter:02} 
with proper and location names extracted from different online 
sources~\footnote{\href{http://alasmaa.net/}{http://alasmaa.net/ }, 
\href{http://ar.wikipedia.org/}{http://ar.wikipedia.org/}}
as well as biblical sources~\footnote{Genesis 4:17-23; 5:1-32; 9:28-10:32; 11:10-32; 25:1-4, 12-18; 36:1-37:2; Exodus 6:14-25; Ruth 4:18-22; 1 Samuel 14:49-51; 1 Chronicles 1:1-9:44; 14:3-7; 24:1; 25:1-27:22; Nehemiah 12:8-26; Matthew 1:1-16; Luke 3:23-38}.

@@

In this paper we make the following contributions. 
\begin{itemize}
\item We automate the segmentation of hadith books into separate narrations and extract chains of narrators 
%into a partial order relation graph 
using a knowledge based approach with morphological analysis.
\item We optimize our analysis through utilizing punctuation and learning previously unknown names.
\item We build from different chains of narrators a partial order relation graph based on merging common narrators.
\item We introduce a specialized distance metric for Arabic names used in Arabic literature. 
\item We highly outperform the currently available narrator extraction tool. 
\item We provide a working and efficient implementation of our work. 
\end{itemize}

The rest of this paper is organized as follows. 
We describe our approach to narrator chain extraction
in Section~\ref{sec:islamic}.
We describe how we build the partial order relation
graph in Section~\ref{sec:graph}.
We discuss related work in Section~\ref{sec:related}.
We present our results in Section~\ref{sec:results} and 
we conclude in Section~\ref{sec:future}.

@@

The process also benefits from name connector 
words and narrator connector words to enhance the
accuracy of the chain extraction and to segment the 
chain accurately into narrators.
Once we meet a name connector or a narrator 
connector word we reset the word counters that we
compare against the thresholds.

(Figure~\ref{f:chain} is confusing since it misses some information uses other symbols like 
those of $\theta$ and is redundant anyways)

@@

\subsection{Optimizations}

(explain better and merged information present about them in "Results" section)

In addition to the controller, 
NaGEMA uses punctuation marks to refine its analysis and 
increase the accuracy of name detection, narration detection and narrator boundaries. 
It also learns narrator names from simple patterns that 
happen within narrator connectors and that are not 
reported by Sarf as names. 
NaGEMA learns names of narrators that consist of words missing from 
the Sarf lexicon.
NaGEMA learns the narrator names based on their context as they occurred 
between narrator connectors or contained name connectors.

\section{The narrator graph}
\label{sec:graph}

@@

The directed acyclic graph (DAG) 
in Figure~\ref{f:narrators} shows a partial order 
relation (POR) between narrators.
We automatically extracted the POR from the three books of 
hadith~\cite{IbnHanbal,AlKulayni,AlTousi}.
The nodes in boxes are the \RL{matn} of the hadith, 
and the other nodes are the narrators.
Given a set of chains of narrators similar to 
Figure~\ref{f:exhadith} we formed the DAG using graph algorithms 
that merged equal names in the chains.

\subsection{Motivation}

The hadith POR is instrumental to automate
localization of narrators in biography documents and
to segment biography documents in future work.
Consider a biography that discusses a narrator.
It mentions his professors and students.
We plan to segment the biographies with a graph 
coloring algorithm
that parses the text looking for names. 
Once a name is found it locates it in the POR graph
and colors the POR node corresponding
to it.
The narrator predecessors in the POR are supposed to be
his professors and the narrator successors in the POR 
are supposed to be his students. 
Consequently, whenever a name moves away from the 
colored local cluster in the POR graph, 
we know that we are within a new biography.

Once the biographies are analyzed, we can annotate
the POR with qualifiers of the narrators that reflect
their authenticity. 
We can also annotate the POR with the locations and 
the time the narrators lived in. 
Then we can check for time and location overlap
to discover inconsistent narrations.
This motivates our interest to automatically generate 
the POR graph.

@@

(modify according to hash)

\begin{table}[tb]
%\centering
%\resizebox{.9\columnwidth}{!}{
\begin{tabular} {p{6cm}}
\begin{Verbatim}[fontsize=\relsize{-2},
frame=topline,framesep=4mm,label=\fbox{Merge Algorithm},
commandchars=\\\{\}, codes={\catcode`$=3\catcode`_=8}]
Merge(Chain $C$[])
  Chain $C_c$[]=\{\}
  foreach Chain $c_1$ in $C$ 
    $C_c = C_c \cup c_1$
    foreach Chain $c_2$ in $C-C_c$
      foreach Narrator $n_1$ in $c_1$
        start = max( $n_1$.index - radius, 0)
        end = min($n_1$.index + radius, $n_2$.size)
        $n_2$ = $c_2$[start]
        while (true)
          if (Distance($n_1$, $n_2$) < $\theta$)
            mergeNodes($n_1$, $n_2$)
          if ($n_2$ = $c_2$[end])
            break
          $n_2$ = $c_2$.next($n_2$)        
\end{Verbatim}
\end{tabular}
%}
\label{t:merge}
\end{table}

The \CodeIn{Merge} Algorithm 
iterates over the chains of narrators looking 
for equivalent narrators and merges the equivalent
narrators.
The \CodeIn{radius} parameter refines the search 
to only consider narrators who belong to the same 
generation. 
For example, a narrator relating directly to the prophet
is definitely not equivalent to a narrator separated
from the prophet by ten or more narrators.

@@

The \CodeIn{Distance} Algorithm 
takes as input two narrators and returns 
their morphological and order distance
from each other. 
The lower the distance is, the more
equivalent the narrators are.

Each narrator has several names and several name and 
narrator connectors. 
\CodeIn{Distance} computes the names of the two narrators
and stores them in $N_1$ and $N_2$ 
and checks whether they are morphologically equivalent. 
If they are not then the distance is set to $\infty$. 
Otherwise the distance is decremented by $\delta_1$. 
Then \CodeIn{Distance} checks whether the morphologically 
equivalent names appear in the same order and 
decrements the distance if so.

\CodeIn{Distance}  pays attention to the $\mathit{NISBA}$ 
words that relate to places and checks on that to make sure
that the two narrators considered equivalent are actually 
from the same place. 
It also checks whether one place such as Baghdad 
is geographically 
contained in the other such as Iraq 
(we skip that in the Algorithm for clarity). 
\CodeIn{Distance} also checks for the equivalency of the
name connectors and the correspondence of the \RL{ibn}
connectors.

The distance metric of NaGEMA performs better than       
existing metrics~\cite{Azmi-2010} as it can detect that  
\RL{`bd al-lah bn `mr} and \RL{ibn `mr} are the same person
while other distance metrics based on simple
string similarity techniques can not.

\subsection{Speedup Optimizations for Large Datasets}

\subsubsection{MultiKey Hashing}

(explain why suitable)

\subsubsection{Three-Level Hierarchy Structure for Graph Nodes}

(possibly omit this info, since is complicated)

%\transfalse
%\novocalize
%\begin{figure*}[tb]
%\center{
%\resizebox{1.8\columnwidth}{!}
%%{ \input{iTreeMohamada}}
%{\includegraphics{iTreeMohamada}}
%\caption{Output of iTree fails to detect \RL{m_hmdA}.}
%\label{f:mohamada}
%}
%\end{figure*}
%\transtrue

The closest work to NaGEMA is 
iTree~\cite{Azmi-2010,iTree}, a 
tool that uses a grammar based approach to solve 
the narrator extraction problem. 
The iTree tool preprocesses the text to remove 
punctuation and diacritics and to normalize white
spaces. 
Then iTree uses a similarity-based approach
to memory based learning~\cite{Azmi-2010} 
in order to perform 
shallow parsing of the preprocessed text. 
The iTree tool uses a parser based on a context
free grammar that enumerates several ways to express
the name of the prophet, several possible words that 
might precede the name of the prophet, several possible
words that might separate two narrators, and several
words that might separate parts of the name of 
the narrators.
The iTree tool fails when a morphological variation of 
one of the 
listed words and not an exact form of it
happens in the text. 
For example, Figure~\ref{f:mohamada} extracted 
using iTree fails to detect \RL{m_hmdA} as a name of the
prophet and does not stop at it when computing the chain 
of the hadith
\transfalse
\setcode{utf8}
%\begin{arabtext}
\RL{
حدثنا وكيع، حدثني سعيد بن السائب، عن داود بن أبي عاصم الثقفي، قال
سألت ابن عمر عن الصلاة، بمنى فقال هل سمعت {\bf محمدا}، صلى الله عليه وسلم قلت
نعم وآمنت فاهتديت به قال فإنه كان يصلي بمنى ركعتين.
%\end{arabtext}
}
\setcode{standard}
\transtrue
\vocalize

%In their grammar they consider a name to be any sequence 
%of 
%legal Arabic characters.

NaGEMA does not need a prepossessing step since 
it uses morphological analysis to process the text. 
While iTree bases its analysis on the patterns 
of words that surround the narrator names, 
NaGEMA bases its analysis on the fact that narrator
names are likely to happen in localities with a high
number of proper names. 
NaGEMA uses the narrator connectors as helpers and
uses morphological analysis to compare the stems of
those words and thus NaGEMA avoids the problem of 
``parser noise words'' that iTree faces.

The iTree tool extracts successfully 86.7 percent of
the narration chains of 90 selected traditions with 
34 simple cases and 56 hard cases. 
We ran NaGEMA against the five hadith that ship with
the tool and the two hadith that are listed in the iTree
paper and NaGEMA was able to extract all narration
chains with no mistakes even on the examples where iTree
reports a ``parser noise'' problem.

As we do not have access to the selected
90 traditions iTree reports on, we claim
that NaGEMA outperforms iTree as it works on
the full text of the presented hadith books
and reports a higher success rate.

Our approach to detect and relate names to verbs
is similar to a local grammar based 
approach~\cite{Traboulsi:09} that detects names by
associating them with verbs 
such as \RL{qaal} ``said'' and \RL{'a_hbar} ``told''. 
The local grammars in~\cite{Traboulsi:09} are also
built into transducers that generate the names and
the learning phase benefits from morphological analysis. 
We differ in that we do not use a local grammar, rather
our tolerance thresholds allow a flexible pattern that is
not easy to capture in a typical regular expression.

Another attempt to Arabic named entity recognition
uses an n-gram approach and maximum 
entropy~\cite{ANERSys:07}. 
The authors use their own corpora and gazetteers to boost 
and train the system and ignore POS tags. 
They achieve major improvements when they use lexical and
morphosyntactic features and a parallel Arabic and English 
corpora to bootstrap noisy features~\cite{Benajiba:2010}.

@@

%\begin{table}[bt]
%\centering
%\caption{Results of the hadith case study with Sarf.}
%%\begin{tabular}{|p{1.5cm}||c|c||c|c||c|c|} \hline
%\resizebox{1.1\columnwidth}{!}{
%\begin{tabular}{lcp{.2cm}cp{.2cm}c} %\cline{2-10}
% &  AlKafi & & AlIstibsar & &IbnHanbal \\ \cline{1-6}
%Word count  &98,943 & & 103,835 & & 20,354 \\ 
% Names       & 12,060  & & 14,613& & 3,013\\
%%Names/Narrator & 1.97 & & 1.84& & 1.25 \\
%Narrators  & 2,623 & & 5,767& & 1,755 \\ 
%%Narrators/Chain  & 4.84 & & 4.76 & &4.05 \\
%Chains  & 542 &  & 1,211& & 433 \\ 
%Ignored names  & 6,400 &  & 3,348 & & 642 \\ \hline
%Segmentation accuracy  & 96\%& & 96\%& & 92\%\\ 
%Chain accuracy & 99\%&  & 99\%& & 97\% \\ 
%%Narrator accuracy  & & & & &  & & & \\ 
%Narrator accuracy  & 91\%& & 90\% & & 90\% \\ \hline
%%Name false positives  & 7\%&  & 4\% & & 4\% \\ \hline
%Running time (secs.)& 1.32 & & 1.31 & & .096\\ \hline 
%\end{tabular}
%}
%\normalsize
%\label{t:hadithresallresults}
%\end{table}
%

\begin{table*}[bt]                                     
\centering                                            
\caption{Results comparing between NaGEMA and iTree.}
\resizebox{2.1\columnwidth}{!}{
\begin{tabular}{p{2.1cm}lp{.1cm}cccp{.1cm}cccp{.1cm}ccc}
%\multicolumn{13}{c}{NaGEMA with all refinements} \\ \cline{3-13}
& & & \multicolumn{3}{c}{segmentation} & &  \multicolumn{3}{c}{sanad boundary}  & &  \multicolumn{3}{c}{name boundary} \\
& & & recall & precision & F-score & & recall & precision & F-score & & recall & precision & F-score \\ \cline{2-14}
\multirow{3}{*}{NaGEMA }&musnad & & 0.999 & 0.999 & 0.999 & & 0.993 & 0.990 & 0.991 & & 0.999 & 0.998 & 0.999 \\
&kafi & & 0.999 & 0.920 & 0.958 & & 0.937 & 0.970 & 0.953 & & 0.993 & 0.988 & 0.990 \\
&istibsar & & 0.999 & 0.973 & 0.986 & & 0.980 & 0.950 & 0.965 & & 0.996 & 0.998 & 0.997 \\ \cline{2-14}
& & & 0.999 & 0.964 & 0.982 & & 0.970 & 0.970 & 0.970 & & 0.996 & 0.995 & 0.995 \\ \hline \hline
%& & &  &  &  & &  &  &  & &  &  & \\
%&\multicolumn{13}{c}{iTree} \\ \cline{3-13}
& & & \multicolumn{3}{c}{accept as hadith} & &  \multicolumn{3}{c}{sanad boundary}  & &  \multicolumn{3}{c}{name boundary} \\
& & & recall & precision & F-score & & recall & precision & F-score & & recall & precision & F-score \\ \cline{2-14}
\multirow{3}{*}{iTree} &musnad & & 0.999 & NA & 0.999 & & 0.997 & 0.977 & 0.987 & & 0.998 & 0.999 & 0.999  \\
&kafi & & 0.043 & NA & 0.083 & & 0.999 & 0.313 & 0.476 & & 0.999 & 0.999 & 0.999 \\
&istibsar & & 0.028 & NA & 0.054 & & 0.875 & 0.583 & 0.700 & & 0.999 & 0.999 & 0.999 \\ \cline{2-14}
& & & 0.357 & NA & 0.379 & & 0.957 & 0.624 & 0.721 & & 0.999 & 0.999 & 0.999 \\ 
 \end{tabular}                 
}                             
\normalsize                   
\label{t:hadithresults} 
\end{table*}

To evaluate our approach, we ran NaGEMA against three books of 
hadith~\cite{IbnHanbal,AlTousi,AlKulayni} and compared the
results to iTree. 
The results are detailed in 
Tables~\ref{t:hadithresults} and \ref{t:nopunctlearnresults}. 
The numbers we present in the tables are the result of a 
manual verification of the correctness of %the outputs of 
10\% of the narrations in the books under consideration.

(discuss annotation and verification process and mention smg about inter-annotator agreement)

We used recall and precision as our evaluation metrics.
We report results on three levels of accuracy: segmentation, sanad boundary and
narrator boundary. 
Segmentation accuracy refers to the success of NaGEMA in
detecting in the plain text of a hadith book the locations of the different
narrations therein.
The sanad boundary accuracy measures the
correctness of individual narrators inside the sanad. 
Finally, the narrator boundary accuracy measures
the correctness of narrator full names.

Precisely defined, segmentation recall refers to the ratio of the narrations correctly detected against the
total number of narrations present. 
Segmentation precision refers to the fraction of correctly detected narration chains compared to all the extracted
narration chains. 
Similarly, sanad boundary recall measures whether we detected all narrators
in the sanad; while sanad boundary precision measures whether we did so without introducing false positives. 
The same concept applies to the narrator boundary where we count the number of words constituting 
a valid narrator.

\subsection{NaGEMA versus iTree}

iTree does not tackle the problem of hadith segmentation. 
In contrast to NaGEMA, it does not work on a book level, 
instead it takes as input an isolated hadith. 
It assumes the hadith book is already segmented.

Consequently, for iTree to extract the narrators from the sanad of a narration, 
the user has to extract and pass the hadith manually.
Still in many cases iTree fails to recognize the structure of
the hadith and does not extract any useful information from the narration. 
This happens for example when a slight variation of the iTree grammar rules 
happens first in the hadith.
In Table~\ref{t:hadithresults}, instead of the segmentation columns 
we report the percentage of times iTree succeeded to recognize the text passed
as a hadith.
This measure is equivalent to the concept of recall; nonetheless, precision
in this context is not applicable since by default the input is required 
to represent correctly extracted narrations.

\subsubsection{On Test-suite Of iTree}

(report on the 90 hadith used in paper of iTree)

\subsubsection{On More Representative Test-suite}

In order to evaluate the approaches used by iTree and NaGEMA without bias,
we augmented the list of names used by 
iTree with those available in the lexicon of Sarf.
%'s lexicon on top of which NaGEMA operates.

First, we compare the recall of the segmentation 
in NaGEMA to that of hadith acceptance rates in iTree. 
Although the task of segmentation is more complex than that of
hadith recognition/acceptance, we still notice that iTree has on the average 35.7\% recall rate compared to 99.9\% for 
NaGEMA. 
However, we notice that the difference in these numbers depended highly on the books under consideration. 
For instance,
for the book of musnad, both approaches correctly detected  practically all narrations under consideration; 
however iTree suffers a lot in 
the books of both Kafi and Istibsar, resulting in a recall of less than 5\%. 
This is due to the less structured nature of the narrations of those books.
For example, iTree's grammar is defined such that a hadith starts by 
an Ikhbar word (\noVocRL{AdaT al-A_hbAar}) such as
\noVocRL{qAl} ``said'', \noVocRL{.hadda_t} ``narrated'', etc...; 
however, the Kafi and Istibsar books followed a flexible style to express the narration 
and most narrations started directly by a narrator name without a preceding Ikhbar word. 
This is a strong evidence of the superiority and tolerance of NaGEMA 
and its threshold based approach
compared to the strict grammar approach used in iTree. 
Clearly, NaGEMA does not fail on such cases, since what it detects as a hadith is based 
on a concentration of names.
Similarly, iTree had a lower sanad boundary precision (62.4\% on the average).
This is mainly due to the grammar that ends a sanad only when a predefined prophet phrase is met.
Prophet phrases are phrases that point to the prophet such as 
\noVocRL{rasowl al-lah} and 
\noVocRL{.sal_A `alayh al-lah wasalam}.

However, especially in the books of Kafi and Istibsar, the prophet is not explicitly mentioned but 
implicitly understood. In those cases, iTree will continue reporting narrators whenever it reaches 
a word that it predefines as a name even if it appears
in the matn of the hadith. 
On the other hand, NaGEMA considers that a sanad ends whenever names become less frequent and measures 
that with a threshold even if it does not reach a prophet
indicative phrase.

Narrator boundary accuracy showed similar results between iTree and 
NaGEMA since we augmented iTree with the names from the lexicon of Sarf.

\subsection{On Unorganized hadith books}

(report results on such hadith books)

\subsection{Effect of punctuation and learning}

\begin{table*}[bt]                                     
\centering                                            
\caption{Results without punctuation and without learning. }
\resizebox{2.1\columnwidth}{!}{
\begin{tabular}{p{2.1cm}lp{.1cm}cccp{.1cm}cccp{.1cm}ccc}
& & & \multicolumn{3}{c}{segmentation} & &  \multicolumn{3}{c}{sanad boundary}  & &  \multicolumn{3}{c}{name boundary} \\
& & & recall & precision & F-score & & recall & precision & F-score & & recall & precision & F-score \\ \cline{2-14}
\multirow{3}{*}{no punctuation} 
& musnad  & & 0.999 & 0.999 & 0.999 & & 0.982 & 0.955 & 0.968 & & 0.999 & 0.998 & 0.999 \\ 
& kafi & & 0.999 & 0.793 & 0.885 & & 0.944 & 0.967 & 0.955 & & 0.996 & 0.938 & 0.966 \\
& istibsar & & 0.999 & 0.999 & 0.999 & & 0.980 & 0.931 & 0.955 & & 0.994 & 0.985 & 0.990 \\  \cline{2-14}
& &  & 0.999 & 0.931 & 0.962 & & 0.969 & 0.951 & 0.960 & & 0.997 & 0.974 & 0.985 \\ \hline \hline
%& & & \multicolumn{3}{c}{segmentation} & &  \multicolumn{3}{c}{sanad boundary}  & &  \multicolumn{3}{c}{name boundary} \\
%& & & recall & precision & F-score & & recall & precision & F-score & & recall & precision & F-score \\ \cline{2-14}
\multirow{3}{*}{no learning}
& musnad & & 0.999 & 0.999 & 0.999 & & 0.894 & 0.985 & 0.937 & & 0.995 & 0.999 & 0.998\\
& kafi & & 0.999 & 0.920 & 0.958 & & 0.937 & 0.970 & 0.953 & & 0.993 & 0.991 & 0.992\\
& istibsar & & 0.999 & 0.973 & 0.986 & & 0.977 & 0.950 & 0.963 & & 0.996 & 0.998 & 0.997\\ \cline{2-14}
 & &  & 0.999 & 0.964 & 0.982 & & 0.936 & 0.968 & 0.951 & & 0.995 & 0.996 & 0.995 \\
 \end{tabular}                 
}                             
\normalsize                   
\label{t:nopunctlearnresults} 
\end{table*}

We are interested in studying the effect of the punctuation 
and the learning optimizations we implemented in NaGEMA.
We utilized punctuation marks to increase the accuracy of narration detection and narrator boundaries. 
We also learned names of narrators that consisted of words missing from 
our lexicon.
We learned the narrator names based on their context as they occurred 
between narrator connectors or contained name connectors.

In Table~\ref{t:nopunctlearnresults}, we report the results of running NaGEMA, 
disabling each of the optimizations at a time.

We notice that punctuations are useful for increasing all precision measures: hadith segmentation, sanad boundary and
narrator boundary. Those measures are enhanced by about 
3\%, 2\%, and 1\% respectively as punctuation related 
information is utilized.  
We observed that punctuation marks helped eliminate confusion that resulted from a 
concentration of names and words that 
can be interpreted falsely as narrator connectors. 
A period or a paragraph mark separated some name concentrations 
dropping each part below the acceptable sanad threshold. 
Punctuation marks also help cleaning the sanad of a narration from 
names that coincidentally appear at the 
end of the matn of a previous narration. 
Names from a previous narration maybe mistaken to belong to the sanad of the current 
narration if they were close enough. 
Punctuation marks reduce such confusions.

Although simple, our name learning optimization enhanced the recall 
of sanad boundaries by about 3.4\% on the average.
This means that we correctly report more narrators.
This is a significant improvement as the task of hadith authentication requires high accuracy. 
In particular, sometimes the presence of a specific narrator in the chain can render a 
hadith as not trustworthy since the narrator is not considered credible.
%That's why it is important not to lose any of the narrators that appear in the sanad.

\subsection{Distance metric}
Finally, the distance metric had 0.993 precision,
0.999 recall and an F-score of 0.997.

(explain why successful on this problem)

(show detailed statistics and compare against that of eNarrator)

\subsection{POR Correctness}

(report recall and precision of nodes missing or those added, bc edges cannot be mistaken other than that)

(report recall and precision on the chainnodes per graphnode, that is all what retains to correctness resulting from algorithm)

\section{Significance \& Extensions -- Named Entity Extraction in Other Contexts}

\subsection{Genealogical Lists in the Bible}

(explain what will be equivalent of NMC and NRC and how it will work

\subsection{Place Names in Travel Itineraries}

(check if it can be applicable)

\subsection{Extending Arabic People Name Gazetteers}

(briefly explain how this specific application has general value through contributing to the general task of entity extraction [learned names])

@@

We presented NaGEMA, a narrator chain extractor
using morphological analysis. 
NaGEMA automatically extracted narrator chains with 
above 90\% accuracy
from three different hadith books and generated 
directed acyclic graphs relating the narrators 
in a partial order relation.

Up to our knowledge, NaGEMA is the first 
tool that automates hadith segmentation and builds
a comprehensive POR graph.
It uses morphological analysis coupled with 
and a knowledge based FSM to  detect the hadith 
sanad.

The partial order graph is instrumental to automate
localization of narrators in biography documents and
segmenting biography documents in the future.
Once the biographies are analyzed, we can annotate
the POR with qualifiers of the narrators that reflect
their authenticity. 
We also plan to 
enhance the accuracy of hadith detection and the 
narrator merging algorithms. 
%We can also annotate the POR with the locations and 
%the time the narrators lived in. 
%Then we can compute a time and location overlap
%check to discover inconsistent narrations.
%We can also perform several interesting checks using 
%this POR on its own such as checking for the effect of
%one narrator on the hadith literature.
